---
title: "Book Examples Chapter 6"
author: "Pablo Bello"
date: "10/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(corrr)
library(rethinking)
library(janitor)
library(ggpubr)
library(dagitty)
theme_set(theme_pubr(border = TRUE))
```

## Berkson's Paradox
```{r}
###--- Berkson's Paradox

# Let's imagine a world where wealth and intelligence are completely independent
world <- 
  tibble(
  wealth = rnorm(1e4),
  intelligence = rnorm(1e4))

# Now let's pick the most clever and intelligent people to be 'celebrities' in this world
celebrities <- 
  world |> 
  mutate(celeb_score = wealth + intelligence) |> 
  arrange(desc(celeb_score)) |> 
  slice_head(prop = .10) # Pick the top 10%

# What's the relationship between intelligence and wealth for celebrities? 
celebrities |> 
  select(-celeb_score) |> 
  correlate()

```

## Multicollinearity

```{r}
###--- Multicollinearity
data(milk)
tbl <- 
  as_tibble(milk) |> 
  clean_names() |> 
  transmute(kcal = standardize(kcal_per_g),
         fat = standardize(perc_fat),
         lac = standardize(perc_lactose)) |> 
  drop_na()


###--- Model 1
model_1 <- alist(
  kcal ~ dnorm(mu,sigma),
  mu <- alpha + beta * fat,
  alpha ~ dnorm(0,.2),
  beta ~ dnorm(0,.5),
  sigma ~ dexp(1)
)

###--- Fit model 1
model_1f <- quap(model_1,data = tbl)
precis(model_1f)


###--- Model 2
model_2 <- alist(
  kcal ~ dnorm(mu,sigma),
  mu <- alpha + beta * lac,
  alpha ~ dnorm(0,.2),
  beta ~ dnorm(0,.5),
  sigma ~ dexp(1)
)

###--- Fit model 2
model_2f <- quap(model_2,data = tbl)
precis(model_2f)


###--- Model 3 (both predictos)
model_3 <- alist(
  kcal ~ dnorm(mu,sigma),
  mu <- alpha + beta_l * lac + beta_f * fat,
  alpha ~ dnorm(0,.2),
  beta_l ~ dnorm(0,.5),
  beta_f ~ dnorm(0,.5),
  sigma ~ dexp(1)
)

###--- Fit model 2
model_3f <- quap(model_3,data = tbl)
precis(model_3f)

```


```{r}
###--- Simulate the effect of multiollinearity on parameter SDs

sim_coll <- function(r, data){
  data <- data |> 
    mutate(x = rnorm(nrow(data),
                     mean = r * fat,
                     sd = sqrt((1 - r^2) * var(fat))))
  
  model <- lm(kcal ~ fat + x, data=data)
  sqrt(diag(vcov(model)))[2] # Std of parameter
  
}

rep_sim_coll <- function(r, n, data){
  std <- replicate(n, sim_coll(r, data))
  mean(std)
}



results <- 
  lapply(seq(0,.99,.01), function(e){
  
  sd <- rep_sim_coll(e,10, tbl)
  tibble(e,sd)
  
}) |> 
  bind_rows()


results |> 
  ggplot(aes(e, sd)) +
  geom_point()
  


```



## Post-treatment Bias

```{r}
set.seed(061295)

###--- Number of plants
N <- 1e4

###--- Initial height
h0 <- rnorm(N,10,2)

###--- Assign treatments and simulate fungus and growth
treatment <- rep(0:1,each=N/2)
fungus <- rbinom(N,size=1,prob=0.5-treatment*0.4)
h1 <- h0 + rnorm(N, 5-3 * fungus)

###--- Compose a clean tibble
tbl <- tibble(h0,h1,treatment,fungus)



###--- A model for the rate of growth
model_1 <- alist(
  h1 ~ dnorm(mu,sigma),
  mu <- h0 * p,
  p ~ dnorm(0,.25),
  sigma ~ dexp(1))

model_1f <- quap(model_1,data = tbl)
precis(model_1f)

###--- Now predict the growth with fungus and treatment
model_2 <- alist(
  h1 ~ dnorm(mu,sigma),
  mu <- h0 * p,
  p <- alpha + treatment * beta_t + fungus * beta_f,
  beta_t ~ dnorm(0,.5),
  beta_f ~ dnorm(0,.5),
  sigma ~ dexp(1),
  alpha ~ dnorm(0,.2))

model_2f <- quap(model_2,data = tbl)
precis(model_2f)

###--- Fungus is a post-treatment variable, and a mediator between treatment and growth
###--- Now run the model without fungus, just treatment
model_3 <- alist(
  h1 ~ dnorm(mu,sigma),
  mu <- h0 * p,
  p <- alpha + treatment * beta_t,
  beta_t ~ dnorm(0,.5),
  sigma ~ dexp(1),
  alpha ~ dnorm(0,.2))

model_3f <- quap(model_3,data = tbl)
precis(model_3f)
```




## D-Separation
```{r}
###---- DAG for the Fungus model 
plant_dag <- dagitty("dag {
                     h0 -> h1
                     f -> h1
                     t -> f
}")

coordinates(plant_dag) <- list( 
  x = c(H_0 = 0, T = 2, F = 1.5, H_1 = 1),
  y = c(H_0 = 0,T = 0,F = 0, H_1 = 0))

drawdag(plant_dag)
```



## Collider Bias

### Simulate collider bias

- Each year, 20 people are born with uniformly distributed happiness values.
- Each year, each person ages one year. Happiness does not change.
- At age 18, individuals can become married. The odds of marriage each year are proportional to an individual's happiness.
- Once married, an individual remains married.
- After age 65, individuals leave the sample (they move to Spain). 


```{r}
N <- 100
years <- 1:(65-18)

tbl <- 
  tibble(
  agents = 1:N,
  h = runif(N,0,1),
  married = 0,
  age = 18)

married <- c()
for(i in years){
  tbl <- 
    tbl |> 
    mutate(age = age + 1) |> 
    rowwise() |> 
    mutate(married_p = rnorm(1,runif(1,0,1)*.5 + h *.5, .2), 
           married = ifelse(married_p > .9,1,married)) 
  
  married[i] <- sum(tbl$married)

}



tibble(years, married) |> 
  ggplot(aes(years,married)) +
  geom_line()

```



```{r}
###--- Regress happiness on age
tbl <- sim_happiness(seed=1977, N_years=1000)
tbl2 <- 
  as_tibble(tbl) |> 
  filter(age > 17) |> 
  mutate(age = (age - 18)/(65-18)) |> 
  mutate(married = married + 1)


###--- Define the  Model
model_1 <- alist(
  happiness ~ dnorm(mu,sigma),
  mu <- a[married] + beta * age,
  a[married] ~ dnorm(0,1),
  beta ~ dnorm(0,2),
  sigma ~ dexp(1)
)

model_1f <- quap(model_1,data = tbl2)
precis(model_1f, depth = 2)
```


### New simulation

(1) P is some function of G and U
(2) C is some function of G, P, and U
(3) G and U are not functions of any other known variables


```{r}
###---- DAG for the simulation 
fam_dag <- dagitty("dag {
                     U [unobserved]
                     G -> P
                     G -> C
                     P -> C
                     U -> C
                     U -> P
}")

coordinates(fam_dag) <- list( 
  x = c(G = 0, P = 1, C = 1, U = 2),
  y = c(G = 0,P = 0,C = 1, U = 0.5))

drawdag(fam_dag)


###----Define the model

N <- 200 # number of grandparent-parent-child triads
b_GP <- 1 # direct effect of G on P
b_GC <- 0 # direct effect of G on C
b_PC <- 1 # direct effect of P on C
b_U <- 2 # direct effect of U on P and C

###--- Simulate the data 
set.seed(061295)
tbl <- 
  tibble(
    U = 2 * rbern(N,.5) - 1,
    G = rnorm(N),
    P = rnorm(N, b_GP * G + b_U * U),
    C = rnorm(N, b_GC * G + b_PC * P + b_U * U)
  )


###--- Estimate the direct effect of G on C

model <- alist(
  C ~ dnorm(mu,sigma),
  mu <- alpha + beta_P * P + beta_G * G,
  beta_P ~ dnorm(0,1),
  beta_G ~ dnorm(0,1),
  alpha ~ dnorm(0,1),
  sigma ~ dexp(1)
)


model_f <- quap(model,data = tbl)
precis(model_f)


```




