---
title: "Chapter 4 Practice"
author: "Pablo Bello"
date: "9/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(tidyverse)
library(ggpubr)
library(rethinking)

theme_set(theme_pubr(border = TRUE))
```

### 4E1 
The likelihood is $y_i \sim \textrm{Normal}(\mu, \sigma)$

### 4E2
There are two $\mu$ and $\sigma$.

### 4E3

$$Pr(\mu,\sigma|y) = \frac{Pr(y|\mu,\sigma) Pr(y)}{\int \int Pr(y | \mu \sigma) Pr(mu) \Pr(\sigma)d\mu d \sigma}$$


### 4E4

In the model definition $\mu_i = \alpha + \beta x_i$ is the linear model. 

### 4E5 
There are 3 parameters.

### 4M1

```{r}
n <- 1e4

# Prior predictions
priors <- 
  tibble(mu = rnorm(n,0,10),
         sigma = rexp(n,1)) |> 
  mutate(pred_h = rnorm(mu,sigma))

# Plot the prior predictions
priors |> 
  ggplot(aes(pred_h)) +
  geom_density() +
  labs(title = "Predicted values for height from priors",
       x = "Predicted height")
  
```


### 4M2
```{r}
model_h <- 
  alist(h ~ dnorm(mu,sigma),
        mu ~ dnorm(0,10),
        sigma ~ dexp(1))
```

### 4M3


$$
\begin{aligned}
y_i &\sim \textrm{Normal}(\mu,\sigma) \\ 
\mu &= \alpha + \beta x_i \\
\alpha &\sim \textrm{Normal}(0,10) \\
\beta &\sim \textrm{Uniform}(0,1) \\
\sigma &\sim \textrm{Exponential}(1)
\end{aligned}
$$

### 4M4
Assuming both height and year are mean centered. We don't expect students to shrink with time so beta can only be positive. 

$$
\begin{aligned}
h_i &\sim \textrm{Normal}(\mu,\sigma) \\ 
\mu &= \alpha + \beta year_i \\
\alpha &= 0 \\
\beta &\sim \textrm{Lognormal}(0,1) \\
\sigma &\sim \textrm{Uniform}(0,30)
\end{aligned}
$$



### 4M5

That is why I would choose a lognormal prior. 

### 4M6

If that information comes from this data I would not use it to inform my priors, simply because they are priors. If it comes from somewhere else I would do $\sigma \sim \textrm{Uniform}(0,8)$. 

### 4M7
```{r}
###--- Load the data
    tbl <- Howell1
    tbl_2 <- 
      as_tibble(tbl) |> 
      filter(age >= 18) 
    
###--- Define the model
  model <- alist(
    height_c ~ dnorm(mu,sigma),
    mu <- beta * weight_c,
    beta ~ dlnorm(0,1),
    sigma ~ dunif(0,30)
   # alpha is not defined because both variables are mean centered
  )
  
  ###--- Fit the model
  model_fit <- 
    quap(model,
      data = tbl_2)
  
```

