---
title: "Practice Chapter 3 SR"
author: "Pablo Bello"
date: "9/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(rethinking)
library(ggpubr)
```


```{r}
###--- Setup
p_grid <-seq(from=0,to=1,length.out=1000)
prior <-rep(1,1000)
likelihood <-dbinom(6,size=9,prob=p_grid)
posterior <-likelihood*prior
posterior <-posterior/sum(posterior)
set.seed(100)
samples <-sample(p_grid,prob=posterior,size=1e4,replace=TRUE)
```


### 3E1. How much posterior probability lies below p = 0.2?

```{r}
sum(samples < .2)/1e4
```


### 3E2. How much posterior probability lies above p = 0.8?

```{r}
sum(samples > .8)/1e4
```



### 3E3. How much posterior probability lies between p = 0.2 and p = 0.8?

```{r}
sum(samples > .2 & samples < .8)/1e4
```

### 3E4. 20% of the posterior probability lies below which value of p?

```{r}
quantile(samples, .2)
```


### 3E5. 20% ofthe posterior probability lies above which value of p?
```{r}
quantile(samples, .8)
```


### 3E6. Which values of p contain the narrowest interval equal to 66% of the posterior probability?

```{r}
HPDI(samples , prob = .66)
```


### 3E7. Which values of p contain 66% of the posterior probability, assuming equal posterior probability both below and above the interval?

```{r}
###--- The hard way

quantile(samples, ((1 - .66)/2)) # lower bound
quantile(samples, ((1 - .66)/2 + .66)) # upper bound


###--- Or the easy way
PI(samples, prob = .66)

```

### 3M1 

```{r, message = FALSE}
###--- Load my function (same as in ch2 script)
source("week_4/helper_functions.R")

###--- Compute grid approximation for W = 8, L = 7 and a flat prior
tbl <- grid_approx(W = 8, L = 15 - 8,grid_n = 1000,plot = FALSE)

head(tbl)
```


### 3M2 

```{r}
samples <- sample(tbl$p_water, prob = tbl$posterior_probability, size = 1e4, replace = TRUE)

HPDI(samples, .9)
```


### 3M3

```{r}
w <- rbinom(1e4, size = 15, prob = samples) # Some predictions 
sum(w == 8)/1e4 # P of 8 waters in 15 tosses
```

### 3M4
```{r}
w <- rbinom(1e4, size = 9, prob = samples) # Some predictions 
sum(w == 6)/1e4 # P of 6 waters in 9 tosses
```


### 3M5

```{r}
###--- Use the step prior now
tbl2 <- grid_approx(W = 8, L = 15 - 8, grid_n = 1000, plot = FALSE, prior_f = "step")
samples2 <- sample(tbl2$p_water, prob = tbl2$posterior_probability, size = 1e4, replace = TRUE)

HPDI(samples2, .9)
```

**With the new priors the 90% HPDI is  narrower.**


```{r}
w2 <- rbinom(1e4, size = 15, prob = samples2) # Some predictions 
sum(w2 == 8)/1e4 # P of 8 waters in 15 tosses
```

**The probability of 8 water in 15 tosses is slightly higher (from ~.15 to ~.16).**

```{r}
w2 <- rbinom(1e4, size = 9, prob = samples2) # Some predictions 
sum(w2 == 6)/1e4 # P of 8 waters in 15 tosses
```

**And the probability of 6 water in 9 tosses is  higher as well (from ~.17 to ~.24).  Which makes sense when we consider that the new priors favor higher P of water.** 

### 3M6

**Can't be done. There is not enough atoms in the universe.**


### 3H1
```{r}
data(homeworkch3)

###--- Number of boys
b <- sum(birth1) + sum(birth2) 
l <- length(birth1) + length(birth2)
tbl <- grid_approx(W = b,L = l - b,grid_n = 1e4,plot = FALSE)

tbl |> 
  filter(posterior_probability == max(posterior_probability)) |> 
  rename(p_boy = p_water)

```

**0.555 is the value for p_boy that maximizes the posterior probability.** 


### 3H2
```{r}
###--- Sample from posterior
samples <- sample(tbl$p_water, prob = tbl$posterior_probability, size = 1e4, replace = TRUE)

###-- 50%, 89% and 97% posterior density interval
HPDI(samples, .5)
HPDI(samples, .89)
HPDI(samples, .97)

```


### 3H3

```{r}
w <- rbinom(1e4, size = 200, prob = samples) # Some predictions 

tibble(w) |> 
  ggplot(aes(w)) +
  geom_density() +
  theme_pubr(border = TRUE) +
  labs(x = "# boys",
       title = "Simulations from posterior of 200 births",
       caption = "Red vertical line indicates observed value from data") +
  geom_vline(aes(xintercept = b), color = "red")

```

**The model seems to fit well. The observed number of male births is the peak or near the peak of the distribution.**

### 3H4

```{r}
w <- rbinom(1e4, size = 100, prob = samples) # Some predictions 

tibble(w) |> 
  ggplot(aes(w)) +
  geom_density() +
  theme_pubr(border = TRUE) +
  labs(x = "# boys",
       title = "Simulations from posterior of 200 births",
       caption = "Red vertical line indicates observed value from data") +
  geom_vline(aes(xintercept = sum(birth1)), color = "red") 

```

**The model is not as good as before. The distribution of predicted number of boys out of 100 births is not centered around the observed value.**


### 3H5
```{r}
girls <- which(birth1 == 0)
n_girls <- length(girls)
boys_2 <- sum(birth2[girls] == 1)


w <- rbinom(1e4, size = n_girls, prob = samples) # Some predictions 

tibble(w) |> 
  ggplot(aes(w)) +
  geom_density() +
  theme_pubr(border = TRUE) +
  labs(x = "# boys",
       title = "Simulations from posterior of 200 births",
       caption = "Red vertical line indicates observed value from data") +
  geom_vline(aes(xintercept = boys_2), color = "red") 

```

**Even worse fit than before. The model is making predictions that look nothing like the data. So we can guess that the births are not independent. There is a negative correlation between the sex at first birth and second birth. Boys tend to follow girls and girls follow boys. Although in real data the correlation seems to be positive.**





