---
title: "Chapter 3 Modern Dive"
author: "Pablo Bello"
date: "9/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 3 

Again, we are going to go through a couple more exercises to reinforce your data wrangling skills. 

As usual, we will begin by loading the `tidyverse` (which is the collection of packages that includes all the verbs you just learned about). 


```{r}
library(tidyverse)
```


Before, we dealt with Olympic athletes but now we are going to get closer to the peak of human achievement: we are going to look at Mario Kart records. 

The data should be loaded from the the `Data` folder in our directory. 


```{r}
mario_kart <- read_csv("Data/world_records.csv")

glimpse(mario_kart)
```



As always, have a look at the data using `glimpse()`. 

## Question 1 

First, let's keep only the records for `Three Lap`. We are interested in whole races after all. 

```{r}
three_laps <- 
  mario_kart |>  
  filter(type == "Three Lap")

```


Anyone worth their salt knows that `"Rainbow Road"` should not even count as a regular track. How would you create a dataset that does not contain any times recorded on that infamous track? Now, save a dataset that only contains the records achieved at Rainbow Road.

```{r}
###-- Filter out Rainbow Road
three_laps |> 
  filter(track != "Rainbow Road")

###--- filter in Ranbow Road
(rain_road <- 
  three_laps |> 
  filter(track != "Rainbow Road")) 

```



## Question 2 

Use `summarize()` to get the average time at Rainbow Road and the standard deviation of the records. 

```{r}
###--- Average and sd of time
rain_road |> 
  summarise(av_time = mean(time),
            sd_time = sd(time))
```

Now, do the same for the dataset that contains all other tracks. 

```{r}
three_laps |> 
  summarise(av_time = mean(time),
            sd_time = sd(time))
```

Notice any differences? 

**On average it takes more time to complete the other tracks, the standard deviation is also higher.**

## Question 3 

It's definitely unfair and uninformative to lump all non-rainbow-road tracks together. So let's use one of the more powerful tricks in the book: the `group_by` + `summarize` combo. I cannot overstate how much you will use this in your daily life as a data analyst. 

So take our dataset that includes only `Three Lap` times and group it by track, then use summarize to count how many different records have been established on each track. 

```{r}
###--- Count number of observations (recods) by track
three_laps |> 
  count(track) |> 
  arrange(desc(n))
```


What's the track that has had the most records established in it? 

**Toad's Turnpike is the track with most records.**
> Hint: the verb `arrange()` might come in handy here.

## Question 4 

We want to now if there are drivers who have multiple records at each track, and how many records they have. This will involve grouping by both driver and track.

```{r}
three_laps |>
  group_by(player, track) |> 
  summarise(n = n()) |> 
  arrange(desc(n))
  
```


Who is the player that has recorded the most records at any one track and what track was it?

**Penev at choco mountain with 26 records.**

## Question 5 

We still haven't looked at the times on each track. Tell me what the average time for each track is using `group_by` and `summarize`.

```{r}
three_laps |>
  group_by(track) |> 
  summarise(av_time = mean(time)) |> 
  arrange(desc(av_time))
  
```

What track has the highest average time?

**Rainbow Road had the highest average time.**


Let me introduce you to a cool trick. The `slice` function takes a certain number of rows inside of each group. For example we can pick the lowest time for each player like this: 

```{r}
three_laps %>% 
  group_by(player) %>% 
  arrange(time) %>% 
  slice(1) %>% 
  head()
```


`head()` is just there to make sure that I only print a small amount of names. 

Now, it's your turn to show me the best time recorded on each track


The result you get using `group_by` and `arrange` contains a lot of information. So let's use `select` so that your out displays only the name of the track and the lowest time recorded there. 

```{r}
three_laps |> 
  group_by(track) |> 
  arrange(time) |> 
  slice(1) |> 
  head() |> 
  select(track,time)
```


## Question 6 

Let's practice creating new variables. I am interested in the records that have held for longer than 100 days. Create a new variable that is a `1` if `record_duration` is higher than 100 or 0 otherwise. 

```{r}
three_laps <- 
  three_laps |> 
  mutate(long_record = ifelse(record_duration > 100,1,0)) 

```

Now, grouping by player, tell me the total amount of long duration records each player holds. What player has the most long duration records? 

```{r}
three_laps |> 
  filter(long_record == 1) |> 
  count(player) |> 
  arrange(desc(n))
  
```

**MR has the most long duration records.**

## Question 7 

We will finish by strengthening our `joining` skills. We have another dataset called `drivers` that tells us where drivers are from. 

Here's how you can read it in. 

```{r}
drivers <- read_csv("Data/drivers.csv")
```



Now use `left_join` to join `three_laps` and `drivers` together. 

```{r}
three_laps <- 
  three_laps |> 
  left_join(drivers)
```

Now, plot a bar chart for how many records each country has. 

```{r}

three_laps |>
  count(nation) |> 
  ggplot(aes(fct_rev(fct_reorder(nation,n)),n)) +
  geom_col() +
  labs(y = "Number of records",
       x = "",
       title = "Number of Records by Country") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5, hjust=1))

```

