---
title: "Week 1 - Chapter 2"
author: "Pablo Bello"
date: "9/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 1

## Question 1

Let's begin by reviewing how to install packages. Try installing three packages called causact, dplyr, and igraph.These will come in handy in our examples below.

```{r, eval = FALSE}
install.packages(c("causact",
                   "dplyr",
                   "igraph"))

```

## Question 2

Load all three packages in the order they were presented above. You should get messages that look something like this.

```{r}
library(causact)
library(dplyr)
library(igraph)
```

It is not wise to ignore these messages; reading them will often save you a lot of time. For instance, R is telling you that igraph is "masking" an object called as_data_frame from the dplyr package. What on earth does that mean?

Well, it turns out that igraph has a function called as_data_frame. It takes a graph and turns it into a dataframe. Now, it just so happens that dplyr has a function with that exact name, which takes objects at turns them into a dataframe.

Run the following code:

```{r, eval = FALSE}
df <- as_data_frame(x = c(1,2,3))
```

```{r}
df <- dplyr::as_data_frame(x = c(1,2,3))
```

```{r}
glimpse(df)
```

Why did this one work? You just learned that you can invoke a specific package function by passing the package name followed by two colons and the function's name. Using this format, use the n_distinct() function in dplyr to count the number of elements in the following collection of numbers:


```{r}
x <- c(5,6,2,7,9,1)

###--- Number of distinct elements in vector x
dplyr::n_distinct(x)
```

In your first call to as_data_frame, R called the function from igraph. That's why it didn't work: it was expecting a graph but instead got a collection of numbers. Can you figure out why R called the function from igraph instead of dplyr?

**Because we loaded igraph after dplyr so it takes priority.**

## Question 3

You might think it is unfair that I made you use an unknown function n_distinct(). Use the ? operator to figure out what the function does and tell me briefly in your own words.

```{r}
?n_distinct()

```

**The function is equivalent to length(unique()), meaning it takes the unique elements of a vector and then looks at how many of them there are.**

## Question 4

The reason we also loaded the causeact package is because it includes some interesting data. As I did above, use the glimpse() function to look at an object called baseballData.

```{r}
glimpse(baseballData)
```

This function gives you a ton of information.

How many rows does the dataset have?

**The dataset has `r nrow(baseballData)` rows.**


How many columns?

**The dataset has `r ncol(baseballData)` columns.**

What type of variable is the Home column? What about HomeScore?

**Home is a factor, and HomeScore an integer.**

## Question 5

Let's play with the data a bit. We can look at the first row of the data by running the following:

```{r}
baseballData[1,]
```

What does one row in the data represent?

**One observation.**

Similarly, we can select certain columns. For instance, if we want the 2nd and 3rd columns, we could do something like this.

```{r}
baseballData[,2:3] |>  head()
```
What do these two columns represent?

**The value of two variables (columns) for all the observations(rows).**

## Question 6

To finish these exercises, let's review how to build our own dataframes. Let's say you have been collecting data on historical hockey goal scorers and you want to build a dataset about the top ten. You have three variables:

* Name
* Total goals
* The year they started playing in the NHL

Here's data:

```{r}
name <-
  c(
    "Wayne Gretzky",
    "Gordie Howe",
    "Jaromir Jagr",
    "Brett Hull",
    "Marcel Dionne",
    "Phil Esposito" ,
    "Mike Gartner",
    "Alex Ovechkin",
    "Mark Messier" ,
    "Steve Yzerman"
  )

goals <- c(894, 801, 766, 741, 731, 717, 708, 700, 694, 692)

year_started <- c(1979, 1946, 1990, 1986, 1971, 1963, 1979, 2005, 1979, 1983)

```

You can put all these elements together in a dataframe by using the tibble() function. It should look something like this:

```{r}
tbl <-tibble(name,goals,year_started)
  
glimpse(tbl)
```


